Community-Centric Innovation

### 2014-03-01 paper.tex
expanding community networks

\section{Community Networks}
Probably the best example of user-centric networking are community networks.
The community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the user needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

@@

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also a means of getting around censorship.
Totalitarian regimes build censorship into the networks to control which information is distributed among the population.
Filtering the information and ordering Internet blackouts have been used as ways for hindering social movements and placate population unrest.

@@

Mesh technologies such as the \emph{quick mesh project} (qmp) are evolving to make it easier for regular citizens to set up a network in a matter of minutes.
Friendly web interfaces for configuration abstract the underlying technological complexity.
In community networks, local people become familiar with the procedures required to set up, maintain and troubleshoot the network.
In case of emergency, the presence of local experts facilitates quick deployments and reparations.

The decentralized nature of community networks greatly helps in fostering organic growth and also in  resisting node failures.
The absence of central points of failure guarantees that two neighbours will be able to connect to each other as long as they keep their nodes running, regardless of what happens to the rest of the network.
Mesh routing protocols, such as BMX6, allow the network to quickly re-configure itself and route packets around failures to keep connectivity.
These protocols are particularly tailored to work in challenging multi-hop wireless conditions.
To take down a community network it is necessary to shut down all the nodes.

@@

This continuous change that forces to keep the equipment firmware updated represents  an additional layer of protection against spying agencies interested in compromising the network.
It is much more difficult to attack up-to-date system than it is to attack systems running old unpatched firmwares.

@@

All participants are invited to contribute, and contributions are evaluated in an informal process for merit or weaknesses.
The participants can progressively build up a reputation by making high-quality contributions and helping others to accomplish their goals.
The organization progressively evolves towards a meritocratic systems in which those participants that consistently contribute to the value of the community earn the respect of their peers and emerge as natural leaders.

Still, this is not a leadership which is attached to the person. 
The leadership is attached to the contributions.
This protects the community against individuals working for purely selfish interests that do not take into the account the goals of the community.

@@

>>>>

\section{Community Networks}
Probably the best example of user-centric networking are community networks.
Community networks are deployed by the same users that use them.
Consequently, they are carefully tailored to satisfy the users' needs.
This needs are not necessarily only about connectivity.
The users might be interested in learning about the network, and in researching  new services, applications and protocols.

@@

Another aspect that can draw the attention to community networks is privacy.
As commercial networks are involved in mass surveillance programs, the citizens see in community networks a way to opt out from privacy-invading campaigns.
Finally, community networks are also a means of getting around censorship.
Totalitarian regimes build censorship into the networks to control which information is distributed among the population.
Filtering the information and Internet blackouts have been used as ways for hindering social movements and placate population unrest.

@@

Mesh technologies such as the \emph{quick mesh project} (qmp) are evolving to make it easier for regular citizens to set up a network in a matter of minutes.
Friendly configuration web interfaces can be used to abstract the underlying technological complexity.
In community networks, local people become familiar with the procedures required to set up, maintain and troubleshoot the network.
In case of emergency, the presence of local experts facilitates quick deployments and reparations.

The decentralized nature of community networks greatly helps in fostering organic growth and also in  resisting node failures.
The absence of central points of failure guarantees that two neighbours will be able to connect to each other as long as they keep their nodes running, regardless of what happens to the rest of the network.
Mesh routing protocols, such as BMX6, allow the network to quickly re-configure itself and to route packets around failures to keep connectivity.
These protocols are particularly tailored to work in challenging multi-hop wireless conditions.
To take down a community network it is necessary to shut down all the nodes.

@@

This continuous change that forces to keep the equipment's firmware updated represents  an additional layer of protection against spying agencies interested in compromising the network.
It is much more difficult to attack up-to-date systems than systems running old unpatched firmwares.

@@

All participants are invited to contribute, and contributions are evaluated in an informal process for merit or weaknesses.
The participants can progressively build up a reputation by making high-quality contributions and helping others to accomplish their goals.
The organization progressively evolves towards a meritocratic systems in which those participants that consistently contribute to the value of the community earn the respect of their peers and emerge as natural leaders.
In any case, this is not a leadership which is attached to the person. 
The leadership is attached to the contributions.

@@

By being part of the community network, the participants agree in altruistically routing traffic for other participants.
Besides connectivity services, the community networks can also offer networked services such as email, mailing lists, HTTP proxy services, FTP, collaborative documents and social platforms.

There are a large number of community networks and they are very heterogeneous.
The use different technologies such as wireless point-to-point links, wireless multi-point to multi-point links, ADSL or fiber.
Even in the case in which two community networks use the same technology, it is still likely that they use different hardware, firmware or protocols.
The sizes also vary greatly and can range from few tenths to tenths of thousands.
The amount of documentation, supporting applications and organizational structure is also very different in different community networks.

Despite the disparity of models for community networks, they also share some aspects in common.
The most important is the involvement of the users in the planning, financing, deployment and maintainance of the networks.

The aspects that community networks share in common are an interest for connectivity and freedom of choice for the users,
